---
title: "Spatial Economics -- Assignment 3"
author:
  - "Gustav Pirich (h11910449)"
  - "Gabriel Konzceny (h11776041)"
date: "April 2, 2024"
output:
  pdf_document:
    toc: true
    includes:
      in_header: !expr file.path("~/Desktop/GITHUB/spatial_econ/helper/wraper_code.tex")
bibliography: references.bib
nocite: '@*'
header-includes:
  - \usepackage{tcolorbox}
  - \usepackage[default]{lato}
  - \usepackage{rotating}
papersize: a4
geometry: margin=2cm
urlcolor: DarkOrchid!65!black
---


```{r, setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(showtext)
showtext_auto()

pacman::p_load(stringi, stringr, stringdist, haven, sf, dplyr, fuzzyjoin, 
               comparator, digest, zoomerjoin, ggplot2, tidyr, ggthemes, viridis, 
               fixest, conleyreg, plm, stargazer, magrittr, tidyverse, tmap, spdep, 
               igraph, generics, knitr, kableExtra, formatR)
```

\vspace{2em}

\begin{tcolorbox}
\centering \itshape The code that was used in compiling the assignment is available on GitHub at \url{https://github.com/gustavpirich/spatial_econ/blob/main/03_assignment/03_assignmnet.Rmd}.
\end{tcolorbox}

\newpage

# Exercise B
```{r, echo = FALSE}
raster_Africa <- read_sf("./data/03_assignment/dataset/raster_Africa.shp")

geoconflict_main <- read_dta("./data/03_assignment/dataset/geoconflict_main.dta")

intersect_coord <- read_dta("./data/03_assignment/dataset/intersect_coord.dta")
```
## Unit of Observation
```{r, echo = FALSE}
area <- st_area(raster_Africa)

raster_Africa %>%
  mutate(area = st_area(geometry))

data <- data.frame(Cells = c("Min", "Mean", "Median", "Max"), Stats = c(min(area), mean(area), median(area), max(area)))

kable(data)
```
The unit of observation are 'subnational' cells in a raster grid of 1 degree of latitude × 1 degree longitude. At the equator this corresponds to a side length of 110 km. The areal extension of the cells varies with the latitude. If you go forther away from the equator, the area of a cell decreases. 
The table shows the area of the cells. The smallest cell has an area of about 97 857 $km^{2}$, the largest 12 364 $km^{2}$. On average the size of the cells is 11 900 $km^{2}$.  Thus the smalles cell is about 20 percent smaller than the largest cell. 
They differ because of the distortions induced by the projection of the surface of the earth, onto a 2 dimensional raster grid. 

## Interpretation of Coefficients

